---
title: "Olympic Games"
format: 
  dashboard:
    orientation: columns
    nav-buttons: [github]
    github: https://github.com/posit-conf-2024/olympicdash
logo: images/olympics-logo.svg
logo-alt: "Olympics logo with multicolored circles."
---

```{r}
#| label: load-packages
library(tidyverse)
library(flextable)

apply_bg_color <- function(x, j, palette) {
  
  numeric_to_color <- function(values, palette) {
    # Normalize the values to a 0-1 scale
    normalized_values = scales::rescale(values)
    # Map the normalized values to colors
    palette_fun = scales::gradient_n_pal(palette)
    colors = palette_fun(normalized_values)
    return(colors)
  }
  
  # Apply background of each individual cell
  for(col_i in j) {
    score_colors = numeric_to_color(x$body$dataset[[col_i]], palette)
    for (row_i in seq_along(score_colors)) {
      x = bg(x, i = row_i, j = col_i, bg = score_colors[row_i])
    }
  }
  
  return(x)
}
```

```{r}
#| label: load-data
olympics_full <- read_csv("data/olympics.csv")
```

```{r}
#| label: prep-data
olympics <- olympics_full |>
  filter(!is.na(medal)) |>
  separate_wider_delim(
    cols = team, 
    names = c("team", "suffix"), 
    delim = "-", 
    too_many = "merge", 
    too_few = "align_start"
  ) |>
  select(-suffix) |>
  mutate(medal = fct_relevel(medal, "Bronze", "Silver", "Gold"))
```

# 😎 Summer Olympics


```{r}
#| label: summer-prep-data

summer_olympics <- olympics |>
  filter(season == "Summer")
```

## Column - Medals by sport and year {width="65%"}

### Row - Medals by sport {height="60%"}

```{r}
#| label: summer-medals-by-sport
#| title: Medals by sport
#| fig-width: 10
#| fig-asp: 0.618
summer_olympics |>
  mutate(
    sport = fct_lump_n(sport, n = 15),
    sport = fct_infreq(sport),
    sport = fct_rev(sport),
    sport = fct_relevel(sport, "Other", after = 0)
  ) |>
  ggplot(aes(y = sport, fill = medal)) +
  geom_bar() +
  guides(fill = guide_legend(reverse = TRUE)) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Medal"
  ) +
  theme_minimal() +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.9, 0.2),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "gray")
  )
```

### Row - Medals by year {height="40%"}

::: {.card title="Medals by year"}

Due to World War II, no olympic game were held in 1940 and 1944.

```{r}
#| label: summer-medals-by-year
#| fig-asp: 0.3
#| fig-width: 10
summer_olympics |>
  count(year, medal) |>
  ggplot(aes(x = year, y = n, color = medal)) +
  geom_point(size = 0.5) +
  geom_line() +
  guides(color = guide_legend(reverse = TRUE)) +
  scale_x_continuous(breaks = seq(1896, 2020, 8)) +
  labs(
    x = "Year",
    y = NULL,
    color = "Medal"
  ) +
  theme_minimal() +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.9, 0.2),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "gray")
  )
```

:::

## Column - Medals by country {width="35%"}

### Row - Value boxes {height="30%"}

```{r}
#| label: summer-calculate-most-medals

summer_most_golds <- summer_olympics |>
  filter(medal == "Gold") |>
  count(team, sort = TRUE) |>
  slice_head(n = 1L)

summer_most_silvers <- summer_olympics |>
  filter(medal == "Silver") |>
  count(team, sort = TRUE) |>
  slice_head(n = 1L)

summer_most_bronzes <- summer_olympics |>
  filter(medal == "Bronze") |>
  count(team, sort = TRUE) |>
  slice_head(n = 1L)
```

::: {.valuebox icon="award-fill" color="#d4af37"}
Most golds:

`{r} summer_most_golds$n`

`{r} summer_most_golds$team`
:::

::: {.valuebox icon="award-fill" color="#c0c0c0"}
Most silvers:

`{r} summer_most_silvers$n`

`{r} summer_most_silvers$team`
:::

::: {.valuebox icon="award-fill" color="#cd7d32"}
Most bronzes:

`{r} summer_most_bronzes$n`

`{r} summer_most_bronzes$team`
:::

### Row - Tabsets of tables {height="70%" .tabset}

::: {.card title="Top 30 total medals"}

Teams sorted in descending order of medals.

```{r}
#| label: summer-top30-total-medals

summer_olympics |>
  count(team, medal) |>
  pivot_wider(
    names_from = medal, 
    values_from = n, 
    values_fill = 0
  ) |>
  transmute(
    Team = team, 
    Gold, 
    Silver,
    Bronze, 
    Total = Bronze + Gold + Silver
  ) |>
  arrange(desc(Total), Team) |>
  slice_head(n = 30L) |>
  flextable() |>
  align(j = "Team") |>
  apply_bg_color(j = c("Bronze", "Silver", "Gold", "Total"),
                 palette = nord::nord("aurora")) |>
  autofit()
```

:::

::: {.card title="Bottom 30 total medals"}

Teams sorted in ascending order of medals.

```{r}
#| label: summer-bottom30-total-medals
summer_olympics |>
  count(team, medal) |>
  pivot_wider(
    names_from = medal, 
    values_from = n, 
    values_fill = 0
  ) |>
  transmute(
    Team = team, 
    Gold, 
    Silver,
    Bronze, 
    Total = Bronze + Gold + Silver
  ) |>
  arrange(Total, Team) |>
  slice_head(n = 30L) |>
  flextable() |>
  align(j = "Team") |>
  apply_bg_color(j = c("Bronze", "Silver", "Gold", "Total"),
                 palette = nord::nord("aurora")) |>
  autofit()
```

:::

# ❄️ Winter Olympic


```{r}
#| label: winter-prep-data

winter_olympics <- olympics |>
  filter(season == "Winter")
```

## Column - Medals by sport and year {width="65%"}

### Row - Medals by sport {height="60%"}

```{r}
#| label: winter-medals-by-sport
#| title: Medals by sport
#| fig-width: 10
#| fig-asp: 0.618
winter_olympics |>
  mutate(
    sport = fct_lump_n(sport, n = 15),
    sport = fct_infreq(sport),
    sport = fct_rev(sport),
    sport = fct_relevel(sport, "Other", after = 0)
  ) |>
  ggplot(aes(y = sport, fill = medal)) +
  geom_bar() +
  guides(fill = guide_legend(reverse = TRUE)) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Medal"
  ) +
  theme_minimal() +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.9, 0.2),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "gray")
  )
```

### Row - Medals by year {height="40%"}

::: {.card title="Medals by year"}

Due to World War II, no olympic game were held in 1940 and 1944.

```{r}
#| label: winter-medals-by-year
#| fig-asp: 0.3
#| fig-width: 10
winter_olympics |>
  count(year, medal) |>
  ggplot(aes(x = year, y = n, color = medal)) +
  geom_point(size = 0.5) +
  geom_line() +
  guides(color = guide_legend(reverse = TRUE)) +
  scale_x_continuous(breaks = seq(1896, 2020, 8)) +
  labs(
    x = "Year",
    y = NULL,
    color = "Medal"
  ) +
  theme_minimal() +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.9, 0.2),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "gray")
  )
```

:::

## Column - Medals by country {width="35%"}

### Row - Value boxes {height="30%"}

```{r}
#| label: winter-calculate-most-medals

winter_most_golds <- winter_olympics |>
  filter(medal == "Gold") |>
  count(team, sort = TRUE) |>
  slice_head(n = 1L)

winter_most_silvers <- winter_olympics |>
  filter(medal == "Silver") |>
  count(team, sort = TRUE) |>
  slice_head(n = 1L)

winter_most_bronzes <- winter_olympics |>
  filter(medal == "Bronze") |>
  count(team, sort = TRUE) |>
  slice_head(n = 1L)
```

::: {.valuebox icon="award-fill" color="#d4af37"}
Most golds:

`{r} winter_most_golds$n`

`{r} winter_most_golds$team`
:::

::: {.valuebox icon="award-fill" color="#c0c0c0"}
Most silvers:

`{r} winter_most_silvers$n`

`{r} winter_most_silvers$team`
:::

::: {.valuebox icon="award-fill" color="#cd7d32"}
Most bronzes:

`{r} winter_most_bronzes$n`

`{r} winter_most_bronzes$team`
:::

### Row - Tabsets of tables {height="70%" .tabset}

::: {.card title="Top 30 total medals"}

Teams sorted in descending order of medals.

```{r}
#| label: winter-top30-total-medals

winter_olympics |>
  count(team, medal) |>
  pivot_wider(
    names_from = medal, 
    values_from = n, 
    values_fill = 0
  ) |>
  transmute(
    Team = team, 
    Gold, 
    Silver,
    Bronze, 
    Total = Bronze + Gold + Silver
  ) |>
  arrange(desc(Total), Team) |>
  slice_head(n = 30L) |>
  flextable() |>
  align(j = "Team") |>
  apply_bg_color(j = c("Bronze", "Silver", "Gold", "Total"),
                 palette = nord::nord("aurora")) |>
  autofit()
```

:::

::: {.card title="Bottom 30 total medals"}

Teams sorted in ascending order of medals.

```{r}
#| label: winter-bottom30-total-medals
winter_olympics |>
  count(team, medal) |>
  pivot_wider(
    names_from = medal, 
    values_from = n, 
    values_fill = 0
  ) |>
  transmute(
    Team = team, 
    Gold, 
    Silver,
    Bronze, 
    Total = Bronze + Gold + Silver
  ) |>
  arrange(Total, Team) |>
  slice_head(n = 30L) |>
  flextable() |>
  align(j = "Team") |>
  apply_bg_color(j = c("Bronze", "Silver", "Gold", "Total"),
                 palette = nord::nord("aurora")) |>
  autofit()
```

:::
